---
title: "Introduction to the package Random Forests"
author: Alexander Kolb, Joel Reiß, Nils Blum, Luke Böhler
date: (to do)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the package Random Forests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will show you how to work with the package in the following steps:

  - Creating a decision tree for regression or classification 
  
  - Applying cost-complexity pruning to a previously generated tree in order to reduce its size and avoid overfitting

  - Using the bagging algorithm to generate multiple bootstrap datasets and corresponding trees for improved predictions

  - Using the random forest algorithm to generate a diverse ensemble of trees to get better predictions
  
To illustrate the functionality, we will use the Ames Housing dataset throughout the examples.

```{r setup}
library(RandomForests)
library(AmesHousing)
ames <- make_ames() # get dataset
```

### Creating a decision tree for regression or classification

To generate a classification-and-regression-tree (CART) on a given dataset you use the function  ``` generate_cart_tree()  ```:

```
cart <- generate_cart_tree( dataSet = ames,
                            properties = 20,
                            n_nodes = 100,
                            mode = "regression",
                            target = "Sale_Price")
```
<br>
After that you can get the prediction for a datapoint of your choice using ``` predict_tree() ``` :

→ checking for the properties the CART algorithm used:
  
  ```
  cart$properties

  #> [1] "Gr_Liv_Area"    "Garage_Cars"    "Garage_Area"    "Total_Bsmt_SF"  "First_Flr_SF"  "TotRms_AbvGrd"  "Year_Built"     "Full_Bath"     
  #> [9] "Year_Remod_Add" "Mas_Vnr_Area"   "Lot_Area"       "Fireplaces"     "Open_Porch_SF"  "Latitude"       "Lot_Frontage"   "Bsmt_Unf_SF"   
  #> [17] "Bsmt_Full_Bath" "Bedroom_AbvGr"  "Second_Flr_SF"  "Longitude"
  ```

→ creating a datpoint with the same properties:
  
  ```
  datapoint <- data.frame(
    Gr_Liv_Area     = 1600,
    Garage_Cars     = 2,
    Garage_Area     = 500,
    Total_Bsmt_SF   = 1100,
    First_Flr_SF    = 1200,
    TotRms_AbvGrd   = 7,
    Year_Built      = 1995,
    Full_Bath       = 2,
    Year_Remod_Add  = 2005,
    Mas_Vnr_Area    = 150,
    Lot_Area        = 8000,
    Fireplaces      = 1,
    Open_Porch_SF   = 120,
    Latitude        = 42.01,
    Lot_Frontage    = 60,
    Bsmt_Unf_SF     = 300,
    Bsmt_Full_Bath  = 1,
    Bedroom_AbvGr   = 3,
    Second_Flr_SF   = 400,
    Longitude       = -93.62
  )
  ```

→ getting a prediction of the target variable for the entered values:
  
  ```
  predict_cart(cart, datapoint)

  #> [1] 181000
  ```
<br> 
You can also check how well your CART generally predicts the target variable by using ```test_cart() ``` with existing data points from your dataset:

→ checking how well the CART performs:

```
test_results <- test_cart(tree = cart,
                          dataPoints = ames[2900:2930,],
                          mode = "regression",
                          target = "Sale_Price")
```
→ the results are presented as the percentage difference between the predicted and the actual value of the target variable:

```
show_results(test_results)
#>          max         min        mean      median 
#>  163.8036810   0.8823529  28.3260395  22.5000000

```

### Cost-complexyity pruning

If the CART was generated using many data points with numerous features, the function ````prune_tree()```` can be used afterwards to reduce the size:

→ number of nodes in the original tree:
```
count_nodes(cart)
#> [1] 199
```
→ cost-complexity pruning:
```
pruned_cart <- prune_tree(cart)
```
→ number of nodes in the pruned tree:
```
plot_cart_tree(pruned_cart$optimal_tree)
count_nodes(pruned_cart$optimal_tree)
#> [1] 45
```
<br>
However, this can slightly reduce the accuracy of the predictions:
```
test_results <- test_cart(tree = pruned_cart$optimal_tree,
                          dataPoints = ames[2900:2930,],
                          mode = "regression",
                          target = "Sale_Price")

show_results(test_results)
#>        max        min       mean     median 
#> 163.803681   3.649273  30.411703  23.562992 
```

### Bagging





  
 












