---
title: "Introduction to the package Random Forests"
author: Alexander Kolb, Joel Reiß, Nils Blum, Luke Böhler
date: 15.03.2026
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the package Random Forests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will show you how to work with the package in the following steps:

  - Creating a decision tree for regression or classification 
  
  - Applying cost-complexity pruning to a previously generated tree in order to reduce its size and avoid overfitting

  - Using the bagging algorithm to generate multiple bootstrap datasets and corresponding trees for improved predictions

  - Using the random forest algorithm to generate a diverse ensemble of trees to get better predictions
  
To illustrate the functionality, we will use the Ames Housing dataset throughout the examples.

```{r setup}
library(RandomForests)
library(AmesHousing)
ames <- make_ames() # get dataset
```

The following datapoint will additionally serve as input for the subsequent prediction functions:

```
datapoint <- data.frame(
  Gr_Liv_Area       = 1710,
  Garage_Cars       = 2,
  Garage_Area       = 548,
  Total_Bsmt_SF     = 856,
  First_Flr_SF      = 856,
  Second_Flr_SF     = 854,
  TotRms_AbvGrd     = 8,
  Year_Built        = 2003,
  Year_Remod_Add    = 2003,
  Mas_Vnr_Area      = 196,
  Lot_Area          = 8450,
  Fireplaces        = 0,
  Open_Porch_SF     = 61,
  Enclosed_Porch    = 120,
  Three_season_porch= 0,
  Screen_Porch      = 80,
  Pool_Area         = 0,
  Wood_Deck_SF      = 0,
  Bsmt_Unf_SF       = 150,
  BsmtFin_SF_1      = 706,
  BsmtFin_SF_2      = 0,
  Bsmt_Full_Bath    = 1,
  Bsmt_Half_Bath    = 1,
  Full_Bath         = 2,
  Half_Bath         = 1,
  Bedroom_AbvGr     = 3,
  Kitchen_AbvGr     = 1,
  Lot_Frontage      = 65,
  Mo_Sold           = 2,
  Year_Sold         = 2008,
  Latitude          = 42.03,
  Longitude         = -93.62,
  Sale_Price        = 208500,
  Misc_Val          = 0,
  Low_Qual_Fin_SF   = 0
);
```
Note: The datapoint you choose for predictions must at least contain the features used by the corresponding algorithms. You can check which features have been used with get_model_features(cart/bagging_trees,random_forest).


### Creating a decision tree for regression or classification

To generate a classification-and-regression-tree (CART) on a given dataset you use the function  ``` generate_cart_tree()  ```:

```
cart <- generate_cart_tree( dataSet = ames,
                            properties = 20,
                            n_nodes = 100,
                            mode = "regression",
                            target = "Sale_Price")
```
<br>
After that you can get the prediction for a datapoint of your choice using ``` predict_tree() ``` :
```
predict_cart(cart, datapoint)
# [1] 146000
```
<br> 
You can also check how well your CART generally predicts the target variable by using ```test_cart() ``` with existing data points from your dataset:

→ checking how well the CART performs:

```
test_results <- test_cart(tree = cart,
                          dataPoints = ames[500:1000,],
                          mode = "regression",
                          target = "Sale_Price")
```
→ the results are presented as the percentage difference between the predicted and the actual value of the target variable:

```
show_results(test_results)
#>       max       min      mean    median 
#> 258.16619   0.00000  24.23485  15.04425 

```

### Cost-complexyity pruning

If the CART was generated using many data points with numerous features, the function ````prune_tree()```` can be used afterwards to reduce the size:

→ number of nodes in the original tree:
```
count_nodes(cart)
#> [1] 199
```
→ cost-complexity pruning:
```
pruned_cart <- prune_tree(cart)
```
→ number of nodes in the pruned tree:
```
count_nodes(pruned_cart$optimal_tree)
#> [1] 45
```
<br>
However, this can slightly reduce the accuracy of the predictions:
```
test_results <- test_cart(tree = pruned_cart$optimal_tree,
                          dataPoints = ames[500:1000,],
                          mode = "regression",
                          target = "Sale_Price")

show_results(test_results)
#>          max          min         mean       median 
#> 229.89493792   0.05057284  26.03053878  15.71189280 

```

### Bagging

To further improve prediction performance, the bagging algorithm can be applied, generating multiple bootstrap datasets that are used to train several decision trees. To do this, simply use the function bagging_greedycart():

```
bagging_trees <- bagging_greedycart(data = ames,
                                    n_bootstrapSamples = 20,
                                    properties = 20,
                                    n_nodes = 200,
                                    mode = "regression",
                                    target = "Sale_Price")
```
<br>
Now you can get a prediction for your datapoint with predict_bagging():
```
predict_bagging(bagging_trees, datapoint, mode = "regression")
#> [1] 188948
```
<br>
You can also run a test here to evaluate the overall performance of the bagging model:
```
test_results <- test_bagging(models = bagging_trees,
                             dataPoints = ames[500:1000,],
                             mode = "regression",
                             target = "Sale_Price")


show_results(test_results)
#>          max          min         mean       median 
#> 219.32822350   0.04685315  20.92268138  14.17593540 
```

### Random Forests
Building upon the bagging approach, the Random Forest algorithm further improves prediction performance by introducing randomness in the selection of features for each split. You can use random_forest() to generate such a forest:
```
randomforest <- random_forest(
    data = ames[1:100, ],
    B = 50,
    m = 8,
    target = "Sale_Price",
    mode = "regression"
)
```
<br>
As with the previous functions, you can obtain a prediction using predict_rf():
```
predict_rf(randomforest,datapoint)
#> [1] 196310
```
<br>
The predictive performance of the random forest can also be measured using a test:
```
test_rf(randomforest, ames[500:1000, ])

#>          max          min         mean       median 
#> 226.44126074   0.04279835  20.06655144  11.62264789
```
